%\newcommand{\code}[1]{\colorbox{lightgray}{\texttt{#1}}}
%\lstinline{snippet}
%http://tex.stackexchange.com/questions/65291/code-snippet-in-text

% This "\code{my code}" can be used to highlight small code snippets in the text like names of variables or methods.
\newcommand{\code}[1]{\textcolor{black}{\texttt{#1}}}

% The "\todo{this still has to be done}" is a command that highlights todos in the text.
\newcommand{\todo}[1]{\textcolor{vscodered}{TODO: \texttt{#1}}}

\newcommand{\red}[1]{\textcolor{red}{#1}}

\usepackage{textcomp}
\newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}


% Text snippets

% Variable Analysis Intro with uncertainty
\newcommand{\varintrosimplewithsam}[1]{This section analyzes the correlated relations of the accident variable \textit{#1}. Groups with an insufficient sample size (see \cref{correlation_uncertainty} are neglected and not considered.}
% Variable Analysis Intro without uncertainty
\newcommand{\varintrosimplewithoutsam}[1]{This section analyzes the correlated relations of the accident variable \textit{#1}.}

% Variable Analysis Intro with uncertainty
\newcommand{\varintrowithsam}[2]{This section analyzes the correlated relations of the accident variable \textit{#2}. Groups with an insufficient sample size (see \cref{correlation_uncertainty} are neglected and not considered. The encoding and description of the variable \textit{#2} is shown in \cref{tbl:#1_dataset_#2}.}
% Variable Analysis Intro without uncertainty
\newcommand{\varintrowithoutsam}[2]{This section analyzes the correlated relations of the accident variable \textit{#2}. The encoding and description of the variable \textit{#2} is shown in \cref{tbl:#1_dataset_#2}.}

% Variable Analysis Intro when a single relation shows no significance
\newcommand{\varintronosigsing}[2]{The Kruskal-Wallis test of \textit{#1} - \textit{#2} results in a $p$-value above the defined $\alpha$-level. Therefore the null hypothesis can't be rejected and no significant difference in \textit{#2} for the relations of \textit{#1} are present.}
% Variable Analysis Intro when multiple relations show no significance
\newcommand{\varintronosigmul}[3]{The Kruskal-Wallis tests of #3 result $p$-values above the defined $\alpha$-level. Therefore the null hypothesis can't be rejected and no significant difference in \textit{#2} for these relations are present.}

% Group Analysis Intro with significant alpha and above .0001
\newcommand{\groupintrosig}[5]{The Kruskal-Wallis test of the relation \textit{#1} - \textit{#2} produces a $p$-value of #3, which is below the defined $\alpha$-level. Therefore the null hypothesis can be rejected, meaning that there are significant differences in \textit{#2} between the groups of \textit{#1}. The significant groups can be identified with a pairwise Wilcoxon $T$-test, which is shown in \cref{tbl:wilcoxon_#4_#5_#1_#2}.}
% Group Analysis Intro with significant alpha and below .0001
\newcommand{\groupintrosigsig}[4]{The Kruskal-Wallis test of the relation \textit{#1} - \textit{#2} produces a $p$-value below .0001, which is far below the defined $\alpha$-level. Therefore the null hypothesis can be rejected, meaning that there are strong significant differences in \textit{#2} between the groups of \textit{#1}. The significant groups can be identified with a pairwise Wilcoxon $T$-test, which is shown in \cref{tbl:wilcoxon_#3_#4_#1_#2}.}
